Enforcing Well-bracketed Control Flow and Stack Encapsulation using Linear Capabilities

### 2017-12-07 fully-abstract-comp/prisc-extended-abstract/linear-well-bracketed-control.tex

\documentclass[sigplan, review]{acmart}

@@

% Copyright
%\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% DOI
% \acmDOI{10.475/123_4}

% ISBN
% \acmISBN{123-4567-24-567/08/06}

%Conference
\acmConference[PriSC]{Workshop on Principles of Secure Compilation}{January 2018}{Los Angeles, California USA} 
\acmYear{2018}
\copyrightyear{2018}

@@

\begin{abstract}
We propose and study a new calling convention that provably enforces well-bracketed control flow and local state encapsulation on a capability machine.
The calling convention is based on linear capabilities, a new type of capabilities that are prevented by the hardware from being duplicated.
In addition to designing and formalising this new calling convention, we also contribute a new way to formalise and prove that it effectively enforces well-bracketed control flow and local state encapsulation, building on the concept of fully abstract compilation.
\end{abstract}

@@

An alternative that does not have this requirement, relies on \emph{linear capabilities}, rather than local ones.
Linear capabilities are (to the best of our knowledge) a new concept, similar to CHERI's local capabilities.\footnote{One author of this abstract is also investigating linear capabilities in another application, also submitted to PriSC.}
The idea is that all capabilities are marked as either regular or linear.
Linear ones are subject to special hardware-enforced restrictions, that prevent them from being duplicated.
For example, when linear capabilities are moved to a new location in registers or memory, they are erased from their old location.
A few special instructions are needed to make this workable, e.g.\ a split instruction that splits a memory capability for a region of memory into disjoint capabilities for two parts of the region and a splice instruction that does the reverse.
This linearity is a powerful tool: when passing a linear capability to untrusted code and receiving the same capability back, we are sure that they cannot have stored a copy of it.
Similarly, when we receive, from untrusted code, a linear capability that addresses memory that we know is only ever addressed by linear capabilities, then we are sure that the untrusted code no longer has access that memory.

>>>>

\documentclass[sigplan,screen]{acmart}

@@

% Copyright
\setcopyright{rightsretained}
\acmDOI{}
\acmISBN{}
\acmConference[PriSC]{Workshop on Principles of Secure Compilation}{2018}{Los Angeles} 
\acmYear{2018}
\copyrightyear{2018}
\settopmatter{printacmref=false}

@@

\begin{abstract}
We propose and study a new calling convention that provably enforces well-bracketed control flow and local state encapsulation on a capability machine.
The calling convention is based on linear capabilities, a type of capabilities that are prevented by the hardware from being duplicated.
In addition to designing and formalising this new calling convention, we also contribute a new way to formalise and prove that it effectively enforces well-bracketed control flow and local state encapsulation, building on the concept of fully abstract compilation.
\end{abstract}

@@

An alternative that does not have this requirement, relies on \emph{linear capabilities}, rather than local ones.
Although details in the literature are scarce, linear capabilities have already been implemented in the SAFE machine~\citep{amorim_verified_2014}\footnote{Private communication with Catalin Hrit\c{c}u.} and considered for implementation in CHERI\footnote{Private communication with Robert Watson.}.
The idea is that all capabilities are marked as either regular or linear.
Linear ones are subject to special hardware-enforced restrictions, that prevent them from being duplicated.
For example, when linear capabilities are moved to a new location in registers or memory, they are erased from their old location.
A few special instructions are needed to make this workable, e.g.\ a split instruction that splits a memory capability for a region of memory into disjoint capabilities for two parts of the region and a splice instruction that does the reverse.
This linearity is a powerful tool: when passing a linear capability to untrusted code and receiving the same capability back, we are sure that they cannot have stored a copy of it.
Similarly, when we receive, from untrusted code, a linear capability that addresses memory that we know is only ever addressed by linear capabilities, then we are sure that the untrusted code no longer has access that memory.
